Verifying Spatial Properties of Array Computations

### 2017-04-11 oopsla17/oopsla17.tex
bug fixes; and dealing with different kinds of control flow in the right way
title and abstract tweak

\begin{document}
\title{Understanding and Verifying the Shape of Stencil Computations}

@@

\maketitle

\begin{abstract}
  Arrays computations are the workhorse of numerical modelling and
  computational science.
  Many practitioners are aware of need to ensure program
  correctness, yet very few of the techniques from the programming
  research community are applied by scientists. We aim to change that
  by providing lightweight verification tools to scientists.
  We focus on one error-prone aspect of numerical code: the 
  data access pattern of arrays. 
  Firstly, we report on a code analysis study on twelve real-world
  computational science code base, identifying common idioms of array
  usage and their spatial properties. This provides much needed data
  on array programming idioms common in science, spanning roughly 1.5
  million lines of code.
  From this data, we 
  designed a lightweight declarative specification language capturing
  the majority of access patterns for arrays. We details its semantics
  and algorithms for checking and inference.
  We detail the design and implementation of a verification tool for
  our specification language, which both checks and infers
  specifications.
  We evaluate our tool on the our previous corpus of scientific code
  and give verification case studies of bugs that could be detected
  with our approach. We find roughly 60,000 targets for specification, showing the vast majority
  of stencil computations indeed have a simple, regular, static shape.
\end{abstract}

 %S ubsequently, Fortran continues to be a
%  popular language choice, due to its focus on arrays and its high
%  performance.

%  In this paper, we propose an alternate approach:
%  a specification language and automated tool for verifying the
%  spatial behaviour of stencils. The specification language is based
%  on the hypothesis that stencil computations read from arrays with a
%  regular, fixed pattern which can be captured by a simple
%  set of abstract combinators. These
%  specifications can be embedded as annotations in source code,
%  against which code can be checked for conformance. For legacy code,
%  spatial specifications can be inferred, providing documentation and
%  aiding future software maintenance. We evaluate our specification
%  language and verification tool against
%  a corpus of numerical Fortran code ($\sim$ 1 million lines) for
%  which we generate $\sim$60,000 specifications, showing the vast majority
%  of stencil computations indeed have a simple, regular, static shape.
%\end{abstract}

@@

>>>>

\title{Verifying Spatial Properties of Array Computations}

@@

\begin{abstract}
  Arrays computations are the workhorse of numerical modelling and
  computational science. However, low-level manipulation of array
  indices is a source of program error.  Many practitioners are aware
  of the need to ensure program correctness, yet very few of the
  techniques from the programming research community are applied by
  scientists. We aim to change that by providing targetted lightweight
  verification techniques for scientific code.  We focus on the all
  too common mistake of array offset errors as a generalisation of
  off-by-one errors.  Firstly, we report on a code analysis study on
  twelve real-world computational science code base, identifying
  common idioms of array usage and their spatial properties. This
  provides much needed data on array programming idioms common in
  scientific code.  From this data, we designed a lightweight
  declarative specification language capturing the majority of array
  access patterns via a small set of combinators. We detail a semantic
  model, and the design and implementation of a verification tool for
  our specification language, which both checks and infers
  specifications.  We evaluate our tool on our corpus of scientific
  code and give verification case studies of bug fixes that are
  detected by our approach. We found roughly 60,000 targets for
  specification across roughly 1.2 million lines of code, showing the
  vast majority of array computations read from arrays in a pattern with
  a simple, regular, static shape.
\end{abstract}

%  We focus on one error-prone aspect of numerical code: the 
%  data access pattern of arrays.

\begin{document}
\maketitle

@@

\begin{itemize}
% (Vars, ...)
\item An assignment to a variable, with an RHS comprising some array
  computation, tends to correspond to a reduction (e.g., calculating
the max value in an array).

\end{itemize}
